---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(fdrat)
```

# fdrat - Faster drat

<!-- badges: start -->
<!-- badges: end -->

## Purpose of fdrat Fork

The `drat::insertPackage()` function has a slow write speed because it is based on `base::file.copy` which has a small write buffer (512 on Window, 8192 on Linux). The performance penalty is significant for cloud-based file shares because the small buffer causes excessive write operations and there is some non-negligible overhead for each operation. The issue is exaggerated even more when the R packages are much larger than typical CRAN package sizes.

A simple solution is to replace base `file.copy` with `fs::file_copy`. Unfortunately, this proposed change was not accepted in the drat package repo ([Issue #127](https://github.com/eddelbuettel/drat/issues/127), [Issue #128](https://github.com/eddelbuettel/drat/issues/128)) so this fork was created to address this issue.

## Benchmarks
- `fdrat::insertPackage` is 20-30% faster when inserting a very large package (`r round(file.size("h2o_3.34.0.3.zip")/1e6, 0)` MB) in local repo

```{r}
repo <- tempdir()
writeLines("<!doctype html><title>empty</title>", con = file.path(repo, "index.html"))
microbenchmark::microbenchmark(
  "fdrat::insertPackage" = {
    fdrat::insertPackage("h2o_3.34.0.3.zip", repo)
  },
  "drat::insertPackage" = {
    drat::insertPackage("h2o_3.34.0.3.zip", repo)
  },
  times = 10L,
  unit = "s",
  control = list(order = "block")
)
```

- `fdrat::insertPackage` is much faster when inserting a large package (`r round(file.size("tidyverse_1.3.1.tar.gz")/1e6, 3)` MB) in cloud-based repo ([SMB file shares in Azure Files](https://docs.microsoft.com/en-us/azure/storage/files/files-smb-protocol?tabs=azure-portal))

```{r}
microbenchmark::microbenchmark(
  "fdrat::insertPackage" = {
    fdrat::insertPackage("tidyverse_1.3.1.tar.gz", "//fdrat.file.core.windows.net/fdrat/drat")
  },
  "drat::insertPackage" = {
    drat::insertPackage("tidyverse_1.3.1.tar.gz", "//fdrat.file.core.windows.net/fdrat/drat")
  },
  times = 10L,
  unit = "s",
  control = list(order = "block")
)
```

- `fdrat::insertPackage` allows inserting very large packages (`r round(file.size("h2o_3.34.0.3.zip")/1e6, 0)` MB) in cloud-based repo which would be prohibitively time consuming using drat ([SMB file shares in Azure Files](https://docs.microsoft.com/en-us/azure/storage/files/files-smb-protocol?tabs=azure-portal))

```{r}
microbenchmark::microbenchmark(
  "fdrat::insertPackage" = {
    fdrat::insertPackage("h2o_3.34.0.3.zip", "//fdrat.file.core.windows.net/fdrat/drat")
  },
  times = 1L,
  unit = "s",
  control = list(order = "block")
)
```


## drat
See https://eddelbuettel.github.io/drat/ for detailed documentation.