---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fdrat - Faster drat

<!-- badges: start -->
<!-- badges: end -->

## Purpose of fdrat Fork

The drat package `insertPackage()` function has a slow write speed because it is based on base `file.copy` which has a small write buffer. This issue is worse on Windows than Linux because the `file.copy` buffer in C is smaller (512 vs 8192), but still may not be much of an issue. However, the performance penalty is significant for cloud-based file shares because there is some non-negligible overhead for each write operation and the small buffer causes excessive write operations. The issue is exaggerated even more when the R packages are much larger than typical CRAN package sizes.

A simple solution is to replace base `file.copy` with `fs::file_copy`. Unfortunately, this proposed change was not accepted in the drat package repo ([Issue #127](https://github.com/eddelbuettel/drat/issues/127), [Issue #128](https://github.com/eddelbuettel/drat/issues/127)) so this fork was created to address this issue.

## Benchmarks
Example of benchmarking transfer speeds to [SMB file shares in Azure Files](https://docs.microsoft.com/en-us/azure/storage/files/files-smb-protocol?tabs=azure-portal) using small 1MB file. 

```{r}
microbenchmark::microbenchmark(
  "fs::file_copy" = {
    fs::file_copy("1MB.bin", "//fdrat.file.core.windows.net/fdrat/1MB.bin", overwrite = TRUE)
  },
  "file.copy" = {
    file.copy("1MB.bin", "//fdrat.file.core.windows.net/fdrat/1MB.bin", overwrite = TRUE)
  },
  times = 10L,
  unit = "s",
  control = list(order = "block")
)
```

## drat
See https://eddelbuettel.github.io/drat/ for detailed documentation.